%*******************************************************
% Acknowledgments
%*******************************************************

\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begingroup

\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Acknowledgements} % Acknowledgements section text
\addcontentsline{toc}{chapter}{\tocEntry{Acknowledgements}}

There are many people who have helped me on this journey and it would be remiss to deny this opportunity to thank each of them.

First and foremost, thank you to both Mark van Rossum and Stefano Panzeri, for their advice and supervision throughout all the work described in this thesis.
I surely could not have done this without either of you.

My thanks also go to Alex Thiele, for his advice concerning my work on perceptual learning (described in \autoref{ch:pl}).
On that note, thank you to Xing Chen, for collecting the electrophysiological data described in \autoref{ch:pl} and, along with Mehdi Sanayei, for helping me to understand it.

Next, thank you to Daniel Zaldivar and Yusuke Murayama, for collecting the electrophysiological data, described in Chapters \ref{ch:lam} and \ref{ch:plam}, and for helping me to understand it.
Thank you to Nikos Logothetis, for supervising the collection of this data and enabling the access of resources at the Max Planck Institute.
Also, thank you to Cesare Magri, for laying the foundations for the analysis described in \autoref{ch:lam}.

To everybody at the University of Edinburgh's Neuroinformatics Doctoral Training Centre, thank you for being such an all-round great community.
There are many of you for whom I have the honourable privilege of calling friends, and I am sure this will not be the last we see of each other.

And finally, last but not certainly not least, thank you to my parents and my sister for offering their continual support and encouragement throughout the last few years, before that, and beyond.

\endgroup
